### 2025-03-26 paper.tex
add paper and test0

%% bare_conf_compsoc.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE Computer
%% Society conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference,compsoc]{IEEEtran}
% Some/most Computer Society conferences require the compsoc mode option,
% but others may want the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a sans serif font rather
% than the serif font used in traditional IEEE formatting and thus the need
% to invoke different subfig.sty package options depending on whether
% compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{HFI in RISC-V}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Dinesh Keserla (dk27833)}
\IEEEauthorblockA{University of Texas at Austin\\
dineshkeserla@utexas.edu}
\and
\IEEEauthorblockN{Dev Patel (dbp748)}
\IEEEauthorblockA{University of Texas at Austin\\
devpatel199@utexas.edu}
\and
\IEEEauthorblockN{Joseph Stanley}
\IEEEauthorblockA{University of Texas at Austin\\
stanley.p.joseph@gmail.com}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page (and note that there is less available width in this regard for
% compsoc conferences compared to traditional conferences), use this
% alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
% TODO
\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}

\subsection{Project Goals}
Hardware-assisted fault isolation (HFI) on RISC-V aims to address two critical needs in modern computing systems: preventing security bugs by isolating untrusted or risky code, and better understanding (and improving) performance overheads that arise from current software-based or OS-based sandboxing mechanisms. By offering a minimal set of architectural extensions for in-process isolation, HFI empowers developers to sandbox components (e.g., libraries, plugins, user-defined functions) without resorting to heavyweight processes or virtual machines. Ultimately, the project’s goal is to demonstrate a secure yet performant solution that seamlessly integrates into existing RISC-V designs.

Several recent high-profile vulnerabilities—such as Heartbleed (CVE-2014-0160) and Spectre (CVE-2017-5753, CVE-2017-5715)—underscore the need for stronger isolation between mutually distrusting code modules. System architects and software developers alike have turned to hardware features for efficient compartmentalization. In this spirit, our HFI design for RISC-V builds on prior work that introduced HFI for x86-64 architectures, enabling fast transitions into and out of “sandbox” contexts while enforcing strong data and control-flow checks.

\subsection{Background}
Although hardware-based isolation solutions can be integrated at multiple levels, our approach focuses on tightening the security of low-level software runtimes responsible for loading, executing, and managing untrusted or semi-trusted code within the same process. Concretely, existing in-process isolation techniques—such as WebAssembly-based solutions—often rely on software-based guard pages or software fault isolation (SFI). These approaches can impose substantial memory footprints and performance overheads, particularly in tight loops or heavily sandboxed server-side environments. Moreover, software guard pages require complex interactions with the operating system’s memory manager (e.g., mprotect calls) that slow down sandbox creation or resizing.

Recent SFI approaches, including WebAssembly’s use of guard regions and JIT-based SFI schemes, have shown that in-process isolation is viable but remains limited by performance bottlenecks and compatibility issues. For instance, large FaaS (Function-as-a-Service) platforms and browser engines isolate tens of thousands of small code modules, resulting in high context-switch overheads or inordinate virtual-address consumption. By consolidating and standardizing architectural support for these needs, HFI reduces reliance on kernel calls, shrinks the required memory footprint, and helps unify existing SFI-based runtimes under a lightweight, hardware-managed sandboxing mechanism.

% TODO add stuff about risc-v and qemu

\subsection{Security/Performance of Software-only Isolation}
Software-only isolation has proven susceptible to exploits that take advantage of subtle boundary checks or speculative execution vulnerabilities. Past disclosures—ranging from the Heartbleed bug (CVE-2014-0160), exposing sensitive in-memory data, to speculative execution attacks like Spectre—highlight that user-level components can become an attack vector if not properly contained. Even robust SFI frameworks, such as those used in browser environments, face challenges from side-channel attacks, especially under cross-sandbox conditions where memory layout or branch predictors leak information.

In contrast, a hardware-assisted approach can unify bounds checking and control-flow constraints at the processor level, closing off out-of-bounds reads or writes before they ever reach caches or translation lookaside buffers. Furthermore, hardware-based mediation of system calls—critical for sandboxing unmodified native libraries—promises to eliminate entire classes of bypass vulnerabilities.

Beyond security, performance remains a central issue. Current software-based guard-page or compiler-instrumented approaches can incur 20–40\% runtime overhead on CPU-intensive code, and ephemeral workloads (common in serverless contexts) may pay a high cost to instantiate and tear down per-sandbox memory. Typical solutions (e.g., mprotect, madvise) require system calls and TLB shootdowns that do not scale gracefully.

On RISC-V, these overheads become particularly relevant for microcontroller-class devices—where every cycle matters—and for large-scale data centers wanting to isolate thousands of workloads concurrently. As prior experiments with x86-based HFI show, carefully designed hardware instructions can reduce context-switch time to near that of a function call, while supporting user-space mechanisms for setting up sandbox memory regions. The challenge is thus to replicate and refine these ideas in a clean-slate RISC-V setting without compromising security or drastically altering the instruction set.

\subsection{Approach}
Our work brings HFI concepts to the RISC-V ecosystem by introducing new “region”-based memory restrictions, a dedicated sandbox mode, and fast transition instructions—all while fitting into RISC-V’s existing privileged architecture. Specifically, we rely on the following techniques:
\begin{enumerate}
    \item In-Process Isolation via Regions. Inspired by previous x86-64 prototypes, the processor tracks a small set of metadata registers—each defining accessible addresses and permissions for code or data. When enabled, any out-of-bounds access or forbidden system call triggers an immediate trap, ensuring robust isolation. These hardware-enforced bounds checks happen in parallel with memory lookups.
    \item Lightweight Sandbox Mode. A single hfi\_enter instruction enables isolation, while hfi\_exit disables it. Each sandbox can optionally redirect system calls, enabling complete interposition if the sandbox is untrusted. The overhead of enter/exit can be on the order of tens of cycles, much lower than classical kernel-based context switches.
    \item User-Space Sandbox Configuration. By placing HFI instructions at the user level, we avoid ring transitions for memory setup. This design helps ephemeral tasks or microservices that create and destroy sandboxes frequently, as reconfiguring regions or resizing memory is a matter of writing to hardware registers, not calling into the OS.
    \item Spectre Mitigations. We adopt optional serialization on transition instructions, ensuring that speculative execution cannot bypass newly set bounds or leak privileged data outside the sandbox. Additional hardware checks prevent malicious code from polluting caches or branch predictors with secrets from other regions.
\end{enumerate}

As demonstrated in prior x86 implementations \cite{HFI}, HFI’s hardware cost is kept minimal, and the main processor pipelines remain largely untouched. This ensures that non-sandboxed code experiences negligible slowdowns, while sandboxed code reaps a significant performance advantage compared to purely software-based solutions.

By systematically combining these hardware instructions, metadata registers, and user-space management strategies, our approach aims to preserve the security guarantees demanded by modern sandboxing frameworks, while also reducing overheads that historically have hindered fine-grained in-process isolation.

\section{Progress}

% conference papers do not normally have an appendix

% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi
% TODO

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{HFI}
Shravan Narayan, Tal Garfinkel, Mohammadkazem Taram, Joey Rudek, Daniel Moghimi, Evan Johnson, Chris Fallin, Anjo Vahldiek-Oberwagner, Michael LeMay, Ravi Sahita, Dean Tullsen, and Deian Stefan. 2023. Going beyond the Limits of SFI: Flexible and Secure Hardware-Assisted In-Process Isolation with HFI. In Proceedings of the 28th ACM International Conference on Architectural Support for Programming Languages and Operating Systems, Volume 3 (ASPLOS 2023). Association for Computing Machinery, New York, NY, USA, 266–281. https://doi.org/10.1145/3582016.3582023

\end{thebibliography}

% that's all folks
\end{document}

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

### 2025-03-26 paper.tex
Implementation via QEMU

@@
 By offering a minimal set of architectural extensions for in-process isolation, HFI empowers developers to sandbox components (e.g., libraries, plugins, user-defined functions) without resorting to heavyweight processes or virtual machines.
-Ultimately, the project’s goal is to demonstrate a secure yet performant solution that seamlessly integrates into existing RISC-V designs.
+Ultimately, the project's goal is to demonstrate a secure yet performant solution that seamlessly integrates into existing RISC-V designs.
 
@@
 These approaches can impose substantial memory footprints and performance overheads, particularly in tight loops or heavily sandboxed server-side environments.
-Moreover, software guard pages require complex interactions with the operating system’s memory manager (e.g., mprotect calls) that slow down sandbox creation or resizing.
+Moreover, software guard pages require complex interactions with the operating system's memory manager (e.g., mprotect calls) that slow down sandbox creation or resizing.
 
-Recent SFI approaches, including WebAssembly’s use of guard regions and JIT-based SFI schemes, have shown that in-process isolation is viable but remains limited by performance bottlenecks and compatibility issues.
+Recent SFI approaches, including WebAssembly's use of guard regions and JIT-based SFI schemes, have shown that in-process isolation is viable but remains limited by performance bottlenecks and compatibility issues.
 For instance, large FaaS (Function-as-a-Service) platforms and browser engines isolate tens of thousands of small code modules, resulting in high context-switch overheads or inordinate virtual-address consumption.
@@
 \subsection{Approach}
-Our work brings HFI concepts to the RISC-V ecosystem by introducing new “region”-based memory restrictions, a dedicated sandbox mode, and fast transition instructions—all while fitting into RISC-V’s existing privileged architecture.
+Our work brings HFI concepts to the RISC-V ecosystem by introducing new “region”-based memory restrictions, a dedicated sandbox mode, and fast transition instructions—all while fitting into RISC-V's existing privileged architecture.
 Specifically, we rely on the following techniques:
@@
 
-As demonstrated in prior x86 implementations \cite{HFI}, HFI’s hardware cost is kept minimal, and the main processor pipelines remain largely untouched.
+As demonstrated in prior x86 implementations \cite{HFI}, HFI's hardware cost is kept minimal, and the main processor pipelines remain largely untouched.
 This ensures that non-sandboxed code experiences negligible slowdowns, while sandboxed code reaps a significant performance advantage compared to purely software-based solutions.
@@
 \section{Progress}
+
+\subsection{Implementation}
+The implementation for HFI in RISC-V is being carried out via Quick Emulator (QEMU). 
+QEMU allows for modifying the semantics to add necessary HFI registers and instructions, allowing for validating HFI's ISA semantics and control-flow traps. 
+Moreover, by implementing in QEMU, RISC-V binaries using the created HFI instructions can be verified for functional correctness.
+
+Unfortunately, this implementation is without problems. 
+QEMU lacks support for simulating microarchitectural features such as branch predictors, speculative execution, cache timing, TLB state, and more.
+Therefore, it should be noted that we do implement the serialization on transition instructions to protect against speculative execution outside of the sandbox, an implementation in QEMU will not be able to verify the side-channel resistance.
+Additionally, verfiying performance is problematic as QEMU does not have proper features for verifying cycles.
+
+A more comprehensive implementation for security and performance in addition to ISA semantics can be considered in gem5 as done in the original HFI paper \cite{HFI}. 
+However, such an implementation is comparatively much slower to QEMU in modification of the libraries and execution. 
+Therefore, as a feasibility metric, an implementation in QEMU is ideal in our circumstances.
 

### 2025-03-26 paper.tex
merge

@@
 \and
-\IEEEauthorblockN{Joseph Stanley}
+\IEEEauthorblockN{Joseph Stanley (jps4455)}
 \IEEEauthorblockA{University of Texas at Austin\\
@@
 
+\section{Related Work}
+
+Several alternate approaches have been developed over the years for fine-grained isolation involving hardware-assisted techniques.
+Some systems have proposed extensions to the page table metadata such as Donky \cite{Donky}, whi
+
 % conference papers do not normally have an appendix
@@
 https://doi.org/10.1145/3582016.3582023
+
+\bibitem{Donky}
+David Schrammel, Samuel Weiser, Stefan Steinegger, Martin Schwarzl, Michael
+Schwarz, Stefan Mangard, and Daniel Gruss. 2020. Donky: Domain Keys–Efficient
+In-Process Isolation for RISC-V and x86. In Proceedings of the USENIX Security
+Symposium (USENIX Security).
+
+\bibitem{CHERI}
+Jonathan Woodruff, Robert N.M. Watson, David Chisnall, Simon W. Moore,
+Jonathan Anderson, Brooks Davis, Ben Laurie, Peter G. Neumann, Robert Norton, and Michael Roe. 2014. The CHERI Capability Model: Revisiting RISC
+in an Age of Risk. SIGARCH Comput. Archit. News 42, 3 (2014). https:
+//doi.org/10.1145/2678373.2665740
+
+\bibitem{IMIX}
 

### 2025-03-26 paper.tex
Current Progress 1

@@
 
+\subsection{Current Progress}
+As of the writing of this paper, we have begun a simple modification to the existing RISC-V in QEMU.
+Our work has started by creating a basic internal register called \textit{hfi\_status}, which serves as a flag bit for verifying whether HFI can be toggled on or off via our newly implemented instructions.
+
 \section{Related Work}
-
 Several alternate approaches have been developed over the years for fine-grained isolation involving hardware-assisted techniques.

### 2025-03-26 paper.tex

--- 2025-03-26 paper.tex
@@
 
-Several alternate approaches have been developed over the years for fine-grained isolation involving hardware-assisted techniques.
-Some systems have proposed extensions to the page table metadata such as Donky \cite{Donky}, whi
+Our work on HFI for RSIC-V builds on prior work that established HFI for x86-64 architectures \cite{HFI}. 
+There are also several alternate approaches have been developed over the years for fine-grained isolation involving hardware-assisted techniques.
+Some systems have proposed extensions to the page table metadata to store a per-sandbox ID checked by hardware, such as Donky \cite{Donky} or IMIX \cite{IMIX}.
+Donky has been implemented for both x86 and RISC-V architectures, while IMIX has only been implemented on x86.
+In addition to this, there are also methods that make use of capability-based addressing, a m, which is seen most prominently in Cheri \cite{CHERI}, which has been implemented for RISC-V.
+While Cheri is a strong security model, it requires extensive modification of the software, and hardware including the OS, ABI, and compiler.
 
@@
 \bibitem{IMIX}
+Tommaso Frassetto, Patrick Jauernig, Christopher Liebchen, and Ahmad-Reza
+Sadeghi. 2018. IMIX:In-Process Memory Isolation EXtension. In Proceedings of
+the USENIX Security Symposium (USENIX Security).
 

### 2025-03-26 paper.tex
Current Progress 2

--- 2025-03-26 paper.tex
@@
 Beyond security, performance remains a central issue.
-Current software-based guard-page or compiler-instrumented approaches can incur 20–40\% runtime overhead on CPU-intensive code, and ephemeral workloads (common in serverless contexts) may pay a high cost to instantiate and tear down per-sandbox memory.
+Current software-based guard-page or compiler-instrumented approaches can incur 20-40\% runtime overhead on CPU-intensive code, and ephemeral workloads (common in serverless contexts) may pay a high cost to instantiate and tear down per-sandbox memory.
 Typical solutions (e.g., mprotect, madvise) require system calls and TLB shootdowns that do not scale gracefully.
@@
 \subsection{Approach}
-Our work brings HFI concepts to the RISC-V ecosystem by introducing new “region”-based memory restrictions, a dedicated sandbox mode, and fast transition instructions—all while fitting into RISC-V's existing privileged architecture.
+Our work brings HFI concepts to the RISC-V ecosystem by introducing new "region"-based memory restrictions, a dedicated sandbox mode, and fast transition instructions—all while fitting into RISC-V's existing privileged architecture.
 Specifically, we rely on the following techniques:
@@
 
-A more comprehensive implementation for security and performance in addition to ISA semantics can be considered in gem5 as done in the original HFI paper \cite{HFI}. 
+A more comprehensive implementation for evaluation of security and performance in addition to ISA semantics can be considered in gem5 as done in the original HFI paper \cite{HFI}. 
 However, such an implementation is comparatively much slower to QEMU in modification of the libraries and execution. 
@@
 \subsection{Current Progress}
-As of the writing of this paper, we have begun a simple modification to the existing RISC-V in QEMU.
+As of the writing of this paper, we have begun a simple modification to the existing RISC-V implementation in QEMU.
 Our work has started by creating a basic internal register called \textit{hfi\_status}, which serves as a flag bit for verifying whether HFI can be toggled on or off via our newly implemented instructions.
+Through the use of QEMU logging we display the value of \textit{hfi\_status} and use it as a validator that the instructions have run.
+
+For our first instruction, we have \textit{hfi\_enter}, which sets \textit{hfi\_status} $=1$ when executed. 
+Additionally, we have \textit{hfi\_exit}, which sets \textit{hfi\_status} $=0$.
+RISC-V contains unmapped opcodes that can be used for the additional of new instructions.
+In order to implement our instruction we have opted to use the CUSTOM0 opcode which is $0001011$.
+We are then able to differentiate between the two functions by modifying the funct3 value of the instructions such that \textit{hfi\_enter} has funct3 $= 000$ and \textit{hfi\_enter} has funct3 $= 001$.
+For now, our implementation of the instructions is that they do not accept any parameters and output no values.
+
+\subsection{Next Steps}
 
@@
 In Proceedings of the 28th ACM International Conference on Architectural Support for Programming Languages and Operating Systems, Volume 3 (ASPLOS 2023).
-Association for Computing Machinery, New York, NY, USA, 266–281.
+Association for Computing Machinery, New York, NY, USA, 266-281.
 https://doi.org/10.1145/3582016.3582023
@@
 David Schrammel, Samuel Weiser, Stefan Steinegger, Martin Schwarzl, Michael
-Schwarz, Stefan Mangard, and Daniel Gruss. 2020. Donky: Domain Keys–Efficient
+Schwarz, Stefan Mangard, and Daniel Gruss. 2020. Donky: Domain Keys-Efficient
 In-Process Isolation for RISC-V and x86. In Proceedings of the USENIX Security

### 2025-03-26 paper.tex

@@
 Some systems have proposed extensions to the page table metadata such as Donky \cite{Donky}, whi
+Our work on HFI for RSIC-V builds on prior work that established HFI for x86-64 architectures \cite{HFI}. 
+There are also several alternate approaches have been developed over the years for fine-grained isolation involving hardware-assisted techniques.
+Some systems have proposed extensions to the page table metadata to store a per-sandbox ID checked by hardware, such as Donky \cite{Donky} or IMIX \cite{IMIX}.
+Donky has been implemented for both x86 and RISC-V architectures, while IMIX has only been implemented on x86.
+In addition to this, there are also methods that make use of capability-based addressing, a m, which is seen most prominently in Cheri \cite{CHERI}, which has been implemented for RISC-V.
+While Cheri is a strong security model, it requires extensive modification of the software, and hardware including the OS, ABI, and compiler.
 

### 2025-03-26 paper.tex
Paper

@@
 \subsection{Next Steps}
+Our next course of action is modifying the instructions to properly implement HFI regions and support for modification of these regions.
+More specifically we will need to create the relevant registers for the regions, update the existing instructions to handle HFI initialization, and create new instructions for managing the regions.
+In addition, during sandboxing, we will need to check with the regions to verify memory accesses in accordance with HFI.
+Moreover, we will need to create new tests to validate the HFI implementation in the various cases as regions are introduced to ensure sandboxing is active as necessary.
 

### 2025-03-27 paper.tex
finished references

--- 2025-03-26 paper.tex
@@
 \section{Related Work}
-Several alternate approaches have been developed over the years for fine-grained isolation involving hardware-assisted techniques.
-Some systems have proposed extensions to the page table metadata such as Donky \cite{Donky}, whi
 Our work on HFI for RSIC-V builds on prior work that established HFI for x86-64 architectures \cite{HFI}. 
@@
 Donky has been implemented for both x86 and RISC-V architectures, while IMIX has only been implemented on x86.
-In addition to this, there are also methods that make use of capability-based addressing, a m, which is seen most prominently in Cheri \cite{CHERI}, which has been implemented for RISC-V.
+In addition to this, there are also methods that make use of capability-based addressing, a scheme for controlling access to memory, which is seen most prominently in Cheri \cite{CHERI}, which has been implemented for RISC-V.
 While Cheri is a strong security model, it requires extensive modification of the software, and hardware including the OS, ABI, and compiler.
@@
 the USENIX Security Symposium (USENIX Security).
+
+\bibitem{QEMU}
+About qemu (no date) About QEMU - QEMU documentation. Available at: https://www.qemu.org/docs/master/about/index.html (Accessed: 26 March 2025).
+
+\bibitem{ResearchGate}
+Performance comparison between models and other simulators We evaluated the performance.
+Retrieved from https://www.researchgate.net/figure/Performance-comparison-between-models-and-other-simulators-We-evaluated-the-performance-fig2-341640101
+
+\bibitem{Opcode}
+Anon. Adding a custom instruction in the RVI subset. Retrieved March 26, 2025 from https://pcotret.gitlab.io/riscv-custom/adding-custom.html#custom-opcodes
 

### 2025-04-29 paper.tex
minor change
paper small change
final paper split up

@@
 % in the abstract
-\begin{abstract}
+% \begin{abstract}
 % TODO
-\end{abstract}
+% \end{abstract}
 
@@
 
-\section{Introduction}
-
-\subsection{Project Goals}
+\section{Project Goals}
 Hardware-assisted fault isolation (HFI) on RISC-V aims to address two critical needs in modern computing systems: preventing security bugs by isolating untrusted or risky code, and better understanding (and improving) performance overheads that arise from current software-based or OS-based sandboxing mechanisms.
@@
 System architects and software developers alike have turned to hardware features for efficient compartmentalization.
-In this spirit, our HFI design for RISC-V builds on prior work that introduced HFI for x86-64 architectures, enabling fast transitions into and out of “sandbox” contexts while enforcing strong data and control-flow checks.
+In this spirit, our HFI design for RISC-V builds on prior work that introduced HFI for x86-64 architectures, enabling fast transitions into and out of "sandbox" contexts while enforcing strong data and control-flow checks.
 
-\subsection{Background}
+\section{Background}
+\subsection{SFI stuff?}
+\subsection{HFI}
 Although hardware-based isolation solutions can be integrated at multiple levels, our approach focuses on tightening the security of low-level software runtimes responsible for loading, executing, and managing untrusted or semi-trusted code within the same process.
@@
 For instance, large FaaS (Function-as-a-Service) platforms and browser engines isolate tens of thousands of small code modules, resulting in high context-switch overheads or inordinate virtual-address consumption.
+To address these issues, Hardware-assisted Fault Isolation (HFI)\cite{HFI} was introduced.
 By consolidating and standardizing architectural support for these needs, HFI reduces reliance on kernel calls, shrinks the required memory footprint, and helps unify existing SFI-based runtimes under a lightweight, hardware-managed sandboxing mechanism.
 
-% TODO add stuff about risc-v and qemu
+Hardware-assisted Fault Isolation (HFI) relies on the following core mechanisms:
+\begin{enumerate}
+    \item In-Process Isolation via Regions.
+The processor tracks a small set of metadata registers—each defining accessible addresses and permissions for code or data.
+When enabled, any out-of-bounds access or forbidden system call triggers an immediate trap, ensuring robust isolation.
+These hardware-enforced bounds checks happen in parallel with memory lookups.
+    \item Lightweight Sandbox Mode.
+A single hfi\_enter instruction enables isolation, while hfi\_exit disables it.
+Each sandbox can optionally redirect system calls, enabling complete interposition if the sandbox is untrusted.
+The overhead of enter/exit can be on the order of tens of cycles, much lower than classical kernel-based context switches.
+    \item User-Space Sandbox Configuration.
+By placing HFI instructions at the user level, ring transitions for memory setup are avoided.
+This design helps ephemeral tasks or microservices that create and destroy sandboxes frequently, as reconfiguring regions or resizing memory is a matter of writing to hardware registers, not calling into the OS.
+    \item Spectre Mitigations.
+HFI includes optional serialization on transition instructions, ensuring that speculative execution cannot bypass newly set bounds or leak privileged data outside the sandbox.
+Additional hardware checks prevent malicious code from polluting caches or branch predictors with secrets from other regions.
+\end{enumerate}
 
-\subsection{Security/Performance of Software-only Isolation}
-Software-only isolation has proven susceptible to exploits that take advantage of subtle boundary checks or speculative execution vulnerabilities.
-Past disclosures—ranging from the Heartbleed bug (CVE-2014-0160), exposing sensitive in-memory data, to speculative execution attacks like Spectre—highlight that user-level components can become an attack vector if not properly contained.
+As demonstrated in prior x86 implementations \cite{HFI}, HFI's hardware cost is kept minimal, and the main processor pipelines remain largely untouched.
+This ensures that non-sandboxed code experiences negligible slowdowns, while sandboxed code reaps a significant performance advantage compared to purely software-based solutions.
+
+\subsection{QEMU}
+The implementation for HFI in RISC-V is being carried out via Quick Emulator (QEMU). 
+QEMU allows for modifying the semantics to add necessary HFI registers and instructions, allowing for validating HFI's ISA semantics and control-flow traps. 
+Moreover, by implementing in QEMU, RISC-V binaries using the created HFI instructions can be verified for functional correctness.
+
+% TODO Maybe move below about qemu downsides to future work or limitations section
+Unfortunately, this implementation is not without problems. 
+QEMU lacks support for simulating microarchitectural features such as branch predictors, speculative execution, cache timing, TLB state, and more.
+Therefore, it should be noted that we do implement the serialization on transition instructions to protect against speculative execution outside of the sandbox, an implementation in QEMU will not be able to verify the side-channel resistance.
+Additionally, verfiying performance is problematic as QEMU does not have proper features for verifying cycles.
+
+A more comprehensive implementation for evaluation of security and performance in addition to ISA semantics can be considered in gem5 as done in the original HFI paper \cite{HFI}. 
+However, such an implementation is comparatively much slower to QEMU in modification of the libraries and execution. 
+Therefore, as a feasibility metric, an implementation in QEMU is ideal in our circumstances.
+
+\section{Security/Performance Currently}
+Software-only isolation approaches face significant security challenges.
 Even robust SFI frameworks, such as those used in browser environments, face challenges from side-channel attacks, especially under cross-sandbox conditions where memory layout or branch predictors leak information.
@@
 
-\subsection{Approach}
-Our work brings HFI concepts to the RISC-V ecosystem by introducing new "region"-based memory restrictions, a dedicated sandbox mode, and fast transition instructions—all while fitting into RISC-V's existing privileged architecture.
-Specifically, we rely on the following techniques:
-\begin{enumerate}
-    \item In-Process Isolation via Regions.
-Inspired by previous x86-64 prototypes, the processor tracks a small set of metadata registers—each defining accessible addresses and permissions for code or data.
-When enabled, any out-of-bounds access or forbidden system call triggers an immediate trap, ensuring robust isolation.
-These hardware-enforced bounds checks happen in parallel with memory lookups.
-    \item Lightweight Sandbox Mode.
-A single hfi\_enter instruction enables isolation, while hfi\_exit disables it.
-Each sandbox can optionally redirect system calls, enabling complete interposition if the sandbox is untrusted.
-The overhead of enter/exit can be on the order of tens of cycles, much lower than classical kernel-based context switches.
-    \item User-Space Sandbox Configuration.
-By placing HFI instructions at the user level, we avoid ring transitions for memory setup.
-This design helps ephemeral tasks or microservices that create and destroy sandboxes frequently, as reconfiguring regions or resizing memory is a matter of writing to hardware registers, not calling into the OS.
-    \item Spectre Mitigations.
-We adopt optional serialization on transition instructions, ensuring that speculative execution cannot bypass newly set bounds or leak privileged data outside the sandbox.
-Additional hardware checks prevent malicious code from polluting caches or branch predictors with secrets from other regions.
-\end{enumerate}
+\section{Approach}
+Our work brings HFI concepts to the RISC-V ecosystem by adapting the existing HFI design while fitting into RISC-V's existing privileged architecture.
+RISC-V presents both unique opportunities and challenges for HFI implementation.
+Its extensible ISA and clean-slate design allow for elegant integration of new instructions, while its simpler microarchitecture requires careful consideration of implementation details.
 
-As demonstrated in prior x86 implementations \cite{HFI}, HFI's hardware cost is kept minimal, and the main processor pipelines remain largely untouched.
-This ensures that non-sandboxed code experiences negligible slowdowns, while sandboxed code reaps a significant performance advantage compared to purely software-based solutions.
+By leveraging RISC-V's custom opcode space, we can integrate the necessary HFI instructions without disrupting compatibility with existing software.
+Our implementation focuses on providing the essential HFI functionality while maintaining the performance benefits that make hardware-assisted isolation valuable.
 
-By systematically combining these hardware instructions, metadata registers, and user-space management strategies, our approach aims to preserve the security guarantees demanded by modern sandboxing frameworks, while also reducing overheads that historically have hindered fine-grained in-process isolation.
+Our approach aims to preserve the security guarantees demanded by modern sandboxing frameworks, while also reducing overheads that historically have hindered fine-grained in-process isolation in the RISC-V ecosystem.
 
-\section{Progress}
+% TODO add the opcode limitation for explicit regions and possible solutions
 
-\subsection{Implementation}
-The implementation for HFI in RISC-V is being carried out via Quick Emulator (QEMU). 
-QEMU allows for modifying the semantics to add necessary HFI registers and instructions, allowing for validating HFI's ISA semantics and control-flow traps. 
-Moreover, by implementing in QEMU, RISC-V binaries using the created HFI instructions can be verified for functional correctness.
+\section{Implementation \& Evaluation}
+% TODO dev
 
-Unfortunately, this implementation is without problems. 
-QEMU lacks support for simulating microarchitectural features such as branch predictors, speculative execution, cache timing, TLB state, and more.
-Therefore, it should be noted that we do implement the serialization on transition instructions to protect against speculative execution outside of the sandbox, an implementation in QEMU will not be able to verify the side-channel resistance.
-Additionally, verfiying performance is problematic as QEMU does not have proper features for verifying cycles.
-
-A more comprehensive implementation for evaluation of security and performance in addition to ISA semantics can be considered in gem5 as done in the original HFI paper \cite{HFI}. 
-However, such an implementation is comparatively much slower to QEMU in modification of the libraries and execution. 
-Therefore, as a feasibility metric, an implementation in QEMU is ideal in our circumstances.
+To begin, we implemented the \textit{hfi\_enter} and \textit{hfi\_exit} instructions. These enabled and disabled HFI, respectively, by setting a control and status (CSR) register that we added.
 
@@
 \section{Related Work}
+% TODO dev flush this out
 Our work on HFI for RSIC-V builds on prior work that established HFI for x86-64 architectures \cite{HFI}. 
@@
 \fi
-% TODO
 
@@
 \bibitem{Opcode}
-Anon. Adding a custom instruction in the RVI subset. Retrieved March 26, 2025 from https://pcotret.gitlab.io/riscv-custom/adding-custom.html#custom-opcodes
+Anon. Adding a custom instruction in the RVI subset. Retrieved March 26, 2025 from https://pcotret.gitlab.io/riscv-custom/adding-custom.htmlcustom-opcodes
 

### 2025-04-30 paper.tex
implementation

@@
 \section{Implementation \& Evaluation}
-% TODO dev
+\subsection{Implementation - Intial Setup}
+Since we were implenting our RISC-V HFI extension in QEMU, we cloned the current open-source QEMU repository and began work there.
+We began by taking time to understand how QEMU works, the purpose of the various repositories, and gaining a better understanding of RISC-V.
+We worked with the assumption that the system being run on always supports HFI but it is an easy addition to add any necessary hardware checks to validate that assumption.\\
+Our work then began by creating a basic internal register called \textit{hfi\_status}, which serves as a flag bit for verifying whether HFI can be toggled on or off via our newly implemented instructions.
+Through the use of QEMU logging we display the value of \textit{hfi\_status} and use it as a validator that the instructions have run.
+For our first instruction, we have \textit{hfi\_enter}, which sets \textit{hfi\_status} $=1$ when executed. 
+Additionally, we have \textit{hfi\_exit}, which sets \textit{hfi\_status} $=0$.
+RISC-V contains unmapped opcodes that can be used for the additional of new instructions.
+In order to implement our instruction we have opted to use the CUSTOM0 opcode which is $0001011$.
+We are then able to differentiate between the two functions by modifying the funct3 value of the instructions such that \textit{hfi\_enter} has funct3 $= 000$ and \textit{hfi\_enter} has funct3 $= 001$.
+Adding the control and status register (CSR) \textit{hfi\_status} and implementing the \textit{hfi\_enter} and \textit{hfi\_exit} gave us basic functionality to enable and disable HFI.\\
 
-To begin, we implemented the \textit{hfi\_enter} and \textit{hfi\_exit} instructions. These enabled and disabled HFI, respectively, by setting a control and status (CSR) register that we added.
+\subsection{Implementation - Implicit Regions}
+After testing whether we can toggle the \textit{hfi\_status} register properly, we moved on to implement implicit regions.
+As aforementioned, implicit regions are used for native, unmodified binaries and an important feature of HFI.
+The original HFI paper had four implicit data regions and two implicit code regions.
+In our implementation, we supported an arbitrary number of data and code regions which were defined by a global variable that could be adjusted as needed.
+We also tested using varying numbers of each region and ensured the HFI logic worked seamlessly with the defined number of regions.
+Implicit regions are defined by a base and a mask.\\
+To store metadata about these regions, we created new structs \textit{HFIImplicitDataRegion} and \textit{HFIImplicitCodeRegion} which we stored in the CPU architecture state in arrays of length based on the number of regions we were supporting.
+To set these regions, we implemented the instruction \textit{hfi\_set\_region\_size} which took in a region number, base address, and mask and configured a new data or code region accordingly.
+Then, to set permissions for this region, we added a new instruction \textit{hfi\_set\_region\_permissions} which took a region number and a 64 bit integer encoding the permissions for that region.
+Since a single call to \textit{hfi\_set\_region\_permissions} will be used for a singular region (differentiated using the code region number), the permission encoding for each of the 3 types of region can also differ.
+Implicit data regions store the enabled bit at bit 7, read bit at bit 6, and write bit at bit 5.
+Implicit code regions store the enabled bit at bit 7 and executable bit at bit 6.\\
+Although implicit regions are used for unmodified native binaries, we are still working with the assumption that the software setting up HFI for the binary is a trustable user who will setup the regions properly.
+However, one check we do insert is for uninitialized regions.
+Originally, if a region wasn't setup properly where \textit{hfi\_set\_region\_size} wasn't called but \textit{hfi\_set\_region\_permissions} was called, a region would still be configured with $0x0$ for the base and mask.
+This was an issue as such a region allowed any code/data addresses to pass through.
+Hence, to prevent this, we set the mask on implicit regions to $0x11111111$ initially which would prevent regions not set up properly to fail HFI checks.\\
+To implement our instruction we have opted to use the CUSTOM0 opcode which is $0001011$.
+We are then able to differentiate between these functions and the two aforementioned functions (\textit{hfi\_enter} and \textit{hfi\_exit}) by modifying the funct3 value of the instructions such that \textit{hfi\_set\_region\_size} has funct3 $= 010$ and \textit{hfi\_set\_region\_permissions} has funct3 $= 011$.
+
+\subsection{Implementation - Explicit Regions}
+Next, we moved on to implement explicit regions by defining a new struct \textit{HFIExpllicitDataRegion} to store metadata about explicit regions.
+These structs were also stored in a new array in the CPU Architecture State of length based on the global variable indicating the number of regions we wanted to support.
+While the original HFI paper had the h-prefix \textit{mov} instruct (\textit{hmov}), we had h-prefix load and store instruction.
+This difference is due to the ISA differences between x86 and RISC-V.
+Hence, for our implementation, we focused on providing support for a single region and implemented the following instructions: \textit{hlb0}, \textit{hlh0}, \textit{hlw0}, \textit{hlbu0}, \textit{hlhu0}, \textit{hlwu0}, \textit{hld0}, \textit{hsb0}, \textit{hsh0}, \textit{hsw0}, and \textit{hsd0}.\\
+For each of the load and store instructions, we would check the whether both the beginning of the access address and end of the address access fall within a given data region (region 0 for the h-prefix and 0-postfix instructions).
+If HFI is enabled and the load or store fall within a properly configured region, then the load or store will be allowed through and proceed to load or store the given address by passing on the access addresses to the corresponding non-HFI load or store instruction.
+If the check fails, HFI would trap and prevent access to the address.\\
+Explicit data regions are configured using a base and a bound.
+To set these regions, we use the same \textit{hfi\_set\_region\_base} and \textit{hfi\_set\_region\_permissions} that we used for implicit regions.
+The value passed for a mask in implicit regions was instead used as the bound value when an explicit region was configured.
+For the permissions, explicit data regions store the enabled bit at bit 7, read bit at bit 6, write bit at bit 5, and large region bit at bit 4.\\
+We tested the implementation using this single explicit data region.
+One obvious limitation is how we can scale this to work for multiple data regions, which can be a challenge due to the opcode limitation discussed in section 4.
+However, based on the feasibility of our implementation for the one data region and overall structure of the implementation, we believe our proposed ideas from section 4 can allow easy extension to support more explicit data regions.\\
+For these new load instructions, we used opcode $0001011$.
+For the new store instructions, we used the opcode $1011011$.
+You can find the funct3 values we used to differentiate between them in Table~\ref{tab:hfi_instructions}.
+
+\begin{table}[]
+\caption{Encodings of HFI instruction implemented}
+\label{tab:hfi_instructions}
+\begin{tabular}{|l|l|l|}
+\hline
+                                         &  opcode   & funct3   \\ \hline
+\textit{hfi\_enter}                      &  0001011  &  000     \\ \hline
+\textit{hfi\_exit}                       &  0001011  &  001     \\ \hline
+\textit{hfi\_set\_region\_size}          &  0001011  &  010     \\ \hline
+\textit{hfi\_set\_region\_permissions}   &  0001011  &  011     \\ \hline
+\textit{hfi\_hlb0}                       &  0101011  &  000     \\ \hline
+\textit{hfi\_hlh0}                       &  0101011  &  001     \\ \hline
+\textit{hfi\_hlw0}                       &  0101011  &  010     \\ \hline
+\textit{hfi\_hlbu0}                      &  0101011  &  100     \\ \hline
+\textit{hfi\_hlhu0}                      &  0101011  &  101     \\ \hline
+\textit{hfi\_hlwu0}                      &  0101011  &  110     \\ \hline
+\textit{hfi\_hld0}                       &  0101011  &  011     \\ \hline
+\textit{hfi\_hsb0}                       &  1011011  &  010     \\ \hline
+\textit{hfi\_hsh0}                       &  1011011  &  011     \\ \hline
+\textit{hfi\_hsw0}                       &  1011011  &  100     \\ \hline
+\textit{hfi\_hsd0}                       &  1011011  &  000     \\ \hline
+\end{tabular}
+\end{table}
 

### 2025-04-30 paper.tex
eval and related works

@@
 RISC-V contains unmapped opcodes that can be used for the additional of new instructions.
-In order to implement our instruction we have opted to use the CUSTOM0 opcode which is $0001011$.
+In order to implement our instruction we have opted to use the free opcode $0001011$.
 We are then able to differentiate between the two functions by modifying the funct3 value of the instructions such that \textit{hfi\_enter} has funct3 $= 000$ and \textit{hfi\_enter} has funct3 $= 001$.
@@
 Hence, to prevent this, we set the mask on implicit regions to $0x11111111$ initially which would prevent regions not set up properly to fail HFI checks.\\
-To implement our instruction we have opted to use the CUSTOM0 opcode which is $0001011$.
+To implement our instruction we have opted to use the free opcode $0001011$.
 We are then able to differentiate between these functions and the two aforementioned functions (\textit{hfi\_enter} and \textit{hfi\_exit}) by modifying the funct3 value of the instructions such that \textit{hfi\_set\_region\_size} has funct3 $= 010$ and \textit{hfi\_set\_region\_permissions} has funct3 $= 011$.
@@
 However, based on the feasibility of our implementation for the one data region and overall structure of the implementation, we believe our proposed ideas from section 4 can allow easy extension to support more explicit data regions.\\
-For these new load instructions, we used opcode $0001011$.
-For the new store instructions, we used the opcode $1011011$.
+For these new load instructions, we used the free opcode $0001011$.
+For the new store instructions, we used the free opcode $1011011$.
 You can find the funct3 values we used to differentiate between them in Table~\ref{tab:hfi_instructions}.
@@
 
-\subsection{Current Progress}
-As of the writing of this paper, we have begun a simple modification to the existing RISC-V implementation in QEMU.
-Our work has started by creating a basic internal register called \textit{hfi\_status}, which serves as a flag bit for verifying whether HFI can be toggled on or off via our newly implemented instructions.
-Through the use of QEMU logging we display the value of \textit{hfi\_status} and use it as a validator that the instructions have run.
+\subsection{Evaluation - Testing}
+To validate our HFI implementation and ensure correct functionality across all features, we developed a comprehensive testing infrastructure.
+Our testing framework consists of a suite of assembly programs written specifically to test different aspects of HFI functionality on RISC-V.
+These tests are compiled using the RISC-V GCC toolchain and executed on our modified QEMU binary.\\
+We use a modular testing approach where each assembly program focuses on testing a specific component or feature of HFI.
+Our test suite covers several key areas of functionality: basic HFI operations that verify the core instructions function correctly and properly toggle the HFI status; region configuration tests that ensure memory regions can be properly defined with appropriate size and permissions; memory access tests that verify operations within configured regions succeed while operations outside them fail appropriately; and explicit region access tests that exercise the h-prefix load/store instructions to validate boundary checking and permission enforcement.
+We also implemented control tests to establish baseline functionality in non-HFI contexts.\\
+These test programs are written in RISC-V assembly and leverage a custom set of macros defined in our includes directory that encode our HFI instructions using the appropriate opcodes and operands.
+For example, our HFI enter macro encodes the instruction with opcode 0x0B and function code 0 along with the specified register operands.
+This approach allows us to write readable assembly that directly tests our custom instructions.\\
+We developed a Makefile-based build system that compiles these assembly programs using the RISC-V toolchain (riscv64-unknown-elf-as and riscv64-unknown-elf-ld).
+The resulting ELF binaries are then executed on our custom-built QEMU with additional debug flags to capture instruction execution and CPU state information.\\
+The test execution outputs are captured in log files, allowing us to analyze instruction execution, register state changes, and any traps or exceptions that occur during testing.
+This approach enables us to verify several critical aspects of the implementation: the HFI instructions execute with the expected opcodes, region configurations are properly stored in CPU state, memory accesses are correctly validated against region bounds, and violations trigger appropriate traps when they should occur.\\
+Through this testing infrastructure, we confirmed the correctness of our implementation across various scenarios including successful sandboxing, proper isolation of memory regions, and correct handling of boundary conditions.
+The modular nature of our test suite allows us to incrementally validate features as they are implemented and regression-test them as the implementation evolves.
 
-For our first instruction, we have \textit{hfi\_enter}, which sets \textit{hfi\_status} $=1$ when executed. 
-Additionally, we have \textit{hfi\_exit}, which sets \textit{hfi\_status} $=0$.
-RISC-V contains unmapped opcodes that can be used for the additional of new instructions.
-In order to implement our instruction we have opted to use the CUSTOM0 opcode which is $0001011$.
-We are then able to differentiate between the two functions by modifying the funct3 value of the instructions such that \textit{hfi\_enter} has funct3 $= 000$ and \textit{hfi\_enter} has funct3 $= 001$.
-For now, our implementation of the instructions is that they do not accept any parameters and output no values.
-
-\subsection{Next Steps}
-Our next course of action is modifying the instructions to properly implement HFI regions and support for modification of these regions.
-More specifically we will need to create the relevant registers for the regions, update the existing instructions to handle HFI initialization, and create new instructions for managing the regions.
-In addition, during sandboxing, we will need to check with the regions to verify memory accesses in accordance with HFI.
-Moreover, we will need to create new tests to validate the HFI implementation in the various cases as regions are introduced to ensure sandboxing is active as necessary.
+\subsection{Evaluation - wasm2c}
+We wanted to additionally add some of the evaluation done by the original HFI paper.
+We attempted to modify wasm2c by building on the work done already.
+The wasm2c code converts wasm files to C source and header and we wanted to modify it to use our modified h-prefix load and store instructions.
+We were unable to complete this step unfortunately due to some issues with building and running the module.
+However, this doesn't detract from the implementation of HFI itself.
+Although it would have been valuable to have used our HFI implementation in wasm2c, our unit tests still comprehensively demonstrate that HFI is feasible and functional in RISC-V.
+Our implementation is able to achieve the sandboxing functionality we wanted and effectively shows that HFI can extend to other architectures beyond x86.
+Hence, our work shows the feasibility of HFI in RISC-V and we encourage evaluation similar to that done in the original HFI paper for future work in this area.
 
 \section{Related Work}
-% TODO dev flush this out
+% TODO dev double check if this correct on the donky and cheri work
 Our work on HFI for RSIC-V builds on prior work that established HFI for x86-64 architectures \cite{HFI}.
+HFI provides a simple ISA extension to support secure, flexible, and efficient in-process isolation, addressing limitations of existing software-based techniques, including runtime overheads, scalability limits, Spectre vulnerability, and compatibility issues.\\
 There are also several alternate approaches have been developed over the years for fine-grained isolation involving hardware-assisted techniques.
 Some systems have proposed extensions to the page table metadata to store a per-sandbox ID checked by hardware, such as Donky \cite{Donky} or IMIX \cite{IMIX}.
-Donky has been implemented for both x86 and RISC-V architectures, while IMIX has only been implemented on x86.
-In addition to this, there are also methods that make use of capability-based addressing, a scheme for controlling access to memory, which is seen most prominently in Cheri \cite{CHERI}, which has been implemented for RISC-V.
-While Cheri is a strong security model, it requires extensive modification of the software, and hardware including the OS, ABI, and compiler.
+Donky is a hardware-software co-design providing strong in-process isolation based on dynamic memory protection domains using hardware-backed protection keys.
+Its implementation on RISC-V uses the reserved top 10 bits of page-table entries for protection keys, supporting 1024 keys per process.
+Donky distinguishes itself by managing protection key policies entirely in userspace via a dedicated control and status register (DKRU).
+Access to the DKRU is protected by a hardware call gate, which leverages the RISC-V N extension.
+This architecture provides substantially stronger security guarantees than systems like Intel MPK, where the policy register is unprivileged, allowing Donky to shield against arbitrary code execution without needing binary inspection or rewriting.
+Donky facilitates fast domain switches entirely in userspace, with minimal hardware additions on RISC-V.
+Donky has been implemented for both x86 (using MPK emulation) and RISC-V architectures, while IMIX has only been implemented on x86.\\
+In addition to this, there are also methods that make use of capability-based addressing, a scheme for controlling access to memory, which is seen most prominently in Cheri \cite{CHERI}.
+While initially developed as a hybrid capability model extending the 64-bit MIPS ISA and demonstrated on FPGA, the CHERI approach is described as not being specific to MIPS and is a prominent capability system relevant to discussions of advanced ISAs, including RISC-V (as referenced in related work comparisons).
+CHERI provides byte-granularity memory protection and enables hardware enforcement of memory safety and fault isolation.
+It utilizes a capability coprocessor and tagged memory, where capabilities are stored in dedicated registers and memory locations, protected by hardware tags and unforgeable manipulation instructions.
+Capabilities are unforgeable and define protection domains based on reachability.
+CHERI aims for a high level of source-code and binary compatibility relative to prior pure capability machines, supporting incremental deployment.
+However, adapting and recompiling application code is generally required to utilize its full protection features, and achieving comprehensive object-level memory safety involves extensive modifications throughout the software stack (OS, ABI, compiler).
+CHERI offers strong fine-grained protection within an address space and supports scalable userspace protection domains.\\
+In summary, while Donky leverages protection keys and a userspace monitor with hardware protection for the policy register, and CHERI introduces capabilities for fine-grained, unforgeable memory references requiring significant software stack changes, our work on HFI for RISC-V focuses on a region-based mechanism, minimal hardware modifications, userspace-only operation, and explicit Spectre mitigation for flexible and efficient in-process isolation supporting both adapted and unmodified code.
 

### 2025-04-30 paper.tex
add to related works

@@
 \section{Related Work}
-% TODO dev double check if this correct on the donky and cheri work
 Our work on HFI for RSIC-V builds on prior work that established HFI for x86-64 architectures \cite{HFI}.
@@
 Donky has been implemented for both x86 (using MPK emulation) and RISC-V architectures, while IMIX has only been implemented on x86.\\
-In addition to this, there are also methods that make use of capability-based addressing, a scheme for controlling access to memory, which is seen most prominently in Cheri \cite{CHERI}.
+IMIX (In-Process Memory Isolation Extension) introduces a minimal x86 ISA extension for isolating sensitive data within a process \cite{IMIX}.
+Unlike many isolation mechanisms tied to a specific defense (e.g., MPK for protection keys or CET for shadow stacks), IMIX provides a mitigation-agnostic hardware primitive for protecting security-critical data.
+IMIX enables memory pages to be marked as isolated using a new permission flag.
+These pages can only be accessed with a special instruction, smov, introduced by IMIX.
+This mechanism allows a wide range of memory-corruption defenses (such as CPI or CFI) to protect their internal metadata from arbitrary access by potentially compromised application code.
+IMIX was prototyped using Intel’s simulation infrastructure, with extensions to LLVM and Linux to support its new instruction semantics.
+The authors demonstrated that IMIX significantly reduces the attack surface for in-process defenses with low overhead and better flexibility than existing solutions like MPK, which are limited by unprivileged key registers and require complex mitigation workarounds to be secure.\\
+In addition to this, there are also methods that make use of capability-based addressing, a scheme for controlling access to memory, which is seen most prominently in CHERI \cite{CHERI}.
 While initially developed as a hybrid capability model extending the 64-bit MIPS ISA and demonstrated on FPGA, the CHERI approach is described as not being specific to MIPS and is a prominent capability system relevant to discussions of advanced ISAs, including RISC-V (as referenced in related work comparisons).
@@
 CHERI offers strong fine-grained protection within an address space and supports scalable userspace protection domains.\\
-In summary, while Donky leverages protection keys and a userspace monitor with hardware protection for the policy register, and CHERI introduces capabilities for fine-grained, unforgeable memory references requiring significant software stack changes, our work on HFI for RISC-V focuses on a region-based mechanism, minimal hardware modifications, userspace-only operation, and explicit Spectre mitigation for flexible and efficient in-process isolation supporting both adapted and unmodified code.
+In summary, while Donky leverages protection keys and a userspace monitor with hardware protection for the policy register, and CHERI introduces capabilities for fine-grained, unforgeable memory references requiring significant software stack changes, IMIX proposes a lightweight ISA extension with minimal hardware cost to securely restrict access to sensitive memory used by diverse in-process defenses.
+Our work on HFI for RISC-V focuses on a region-based mechanism, minimal hardware modifications, userspace-only operation, and explicit Spectre mitigation for flexible and efficient in-process isolation supporting both adapted and unmodified code.
 

### 2025-04-30 paper.tex
SFI + HFI + QEMU + Future Work

@@
 \section{Background}
-\subsection{SFI stuff?}
-\subsection{HFI}
-Although hardware-based isolation solutions can be integrated at multiple levels, our approach focuses on tightening the security of low-level software runtimes responsible for loading, executing, and managing untrusted or semi-trusted code within the same process.
-Concretely, existing in-process isolation techniques—such as WebAssembly-based solutions—often rely on software-based guard pages or software fault isolation (SFI).
-These approaches can impose substantial memory footprints and performance overheads, particularly in tight loops or heavily sandboxed server-side environments.
-Moreover, software guard pages require complex interactions with the operating system's memory manager (e.g., mprotect calls) that slow down sandbox creation or resizing.
+\subsection{Software-based Fault Isolation}
+Modern applications consistently rely on a large number of various libraries for key functionality, however, this leads to a problem as bugs within these libraries can become a vector of attack for the entire application.
+To mitigate this, it is critical to contain the scope of what each library can access or affect.
+Traditionally, general isolation was considered at the granularity of processes, containers, and virtual machines, which do achieve the functionality of isolation of libraries through hardware-level protection via the MMU and user-kernel privileges.
+Although such these isolation methods are functional, they have significant performance overhead from the context switching and memory management. 
+Due to the fact that the executing libraries are quite small, it motivates isolation that is fine-grained and lightweight.
 
+These issues are remedied by Software-based Fault Isolation (SFI), which performs isolation in userspace via compiler instrumentation.
+SFI creates a sandbox, which is a restricted execution environment where untrusted code is contained in regards to both memory safety and control-flow integrity.
+Memory accesses are restricted via bounds checks or address masking, while control-flow integrity is maintained by constraining jumps to valid, pre-approved destinations.
+Additionally, to prevent unauthorized interaction with the operating system, system calls from sandboxed code are redirected to a trusted runtime, which will validate and mediate them before execution.
+
+This process of fine-grained isolation within a single address space of the application is referred as \textit{in-process isolation.}
 Recent SFI approaches, including WebAssembly's use of guard regions and JIT-based SFI schemes, have shown that in-process isolation is viable but remains limited by performance bottlenecks and compatibility issues.
 For instance, large FaaS (Function-as-a-Service) platforms and browser engines isolate tens of thousands of small code modules, resulting in high context-switch overheads or inordinate virtual-address consumption.
-To address these issues, Hardware-assisted Fault Isolation (HFI)\cite{HFI} was introduced.
+More explicitly, SFI systems face the following limitations:
+\begin{enumerate}
+  \item Performance overhead: SFI implementations can impose significant runtime overhead with over $40\%$ in practical workloads.
+This stems from the cost of inserting and executing additional instructions for validating memory loads and stores, control-flow validation, and system call interposition.
+\\
+  \item Limited scalability: SFI schemes rely on allocating large, sparsely-mapped virtual memory regions with guard pages (8 GiB per sandbox in WebAssembly) in order to validate memory with the much more performant masking as opposed to base and bound checks.
+This consumes vast amounts of virtual address space, making it impractical to host many concurrent sandboxes.
+\\
+  \item Poor compatibility: SFI requires code to be compiled or transformed in specific ways, which limits its compatibility with unmodified native binaries, low-level system libraries, and code that performs dynamic memory management or emits code at runtime.
+\\
+  \item Vulnerability to spectre attacks: SFI checks are instrumented as checks within the code and like any other code, modern processors may speculatively execute instructions past these checks, potentially leaking sensitive data through a side channel.
+\end{enumerate}
+
+Despite these limitations, SFI is increasingly adopted due to its relatively low overhead compared to traditional isolation mechanisms.
+
+\subsection{HFI: Hardware-based Fault Isolation}
+To address the aforementioned limitations found in SFI, Hardware-assisted Fault Isolation (HFI)\cite{HFI} was introduced.
 By consolidating and standardizing architectural support for these needs, HFI reduces reliance on kernel calls, shrinks the required memory footprint, and helps unify existing SFI-based runtimes under a lightweight, hardware-managed sandboxing mechanism.
 
+Although hardware-based isolation solutions can be integrated at multiple levels, the approach mentioned in \cite{HFI} focuses on tightening the security of low-level software runtimes responsible for loading, executing, and managing untrusted or semi-trusted code within the same process.
 Hardware-assisted Fault Isolation (HFI) relies on the following core mechanisms:
 \begin{enumerate}
-    \item In-Process Isolation via Regions.
-The processor tracks a small set of metadata registers—each defining accessible addresses and permissions for code or data.
+    \item In-Process Isolation via Regions: The processor tracks a small set of metadata registers—each defining accessible addresses and permissions for code or data.
 When enabled, any out-of-bounds access or forbidden system call triggers an immediate trap, ensuring robust isolation.
 These hardware-enforced bounds checks happen in parallel with memory lookups.
-    \item Lightweight Sandbox Mode.
-A single hfi\_enter instruction enables isolation, while hfi\_exit disables it.
+\\
+    \item Lightweight Sandbox Mode: A single hfi\_enter instruction enables isolation, while hfi\_exit disables it.
 Each sandbox can optionally redirect system calls, enabling complete interposition if the sandbox is untrusted.
 The overhead of enter/exit can be on the order of tens of cycles, much lower than classical kernel-based context switches.
-    \item User-Space Sandbox Configuration.
-By placing HFI instructions at the user level, ring transitions for memory setup are avoided.
+\\
+    \item User-Space Sandbox Configuration: By placing HFI instructions at the user level, ring transitions for memory setup are avoided.
 This design helps ephemeral tasks or microservices that create and destroy sandboxes frequently, as reconfiguring regions or resizing memory is a matter of writing to hardware registers, not calling into the OS.
-    \item Spectre Mitigations.
-HFI includes optional serialization on transition instructions, ensuring that speculative execution cannot bypass newly set bounds or leak privileged data outside the sandbox.
+\\
+    \item Spectre Mitigations: HFI includes optional serialization on transition instructions, ensuring that speculative execution cannot bypass newly set bounds or leak privileged data outside the sandbox.
 Additional hardware checks prevent malicious code from polluting caches or branch predictors with secrets from other regions.
+\\
 \end{enumerate}
@@
 \subsection{QEMU}
-The implementation for HFI in RISC-V is being carried out via Quick Emulator (QEMU). 
-QEMU allows for modifying the semantics to add necessary HFI registers and instructions, allowing for validating HFI's ISA semantics and control-flow traps. 
+
+Quick Emulator (QEMU) is a free and open-source emulator for computer processors of varying architecture. 
+QEMU is able to emulate various popular architectures such as x86, ARM, RISC-V, and more.
+The core functionality of emulating various CPU architectures is achieved through \textit{dynamic binary translation}, in which guest machine instructions are converted to equivalent host machine instructions at runtime.
+This allows it to execute guest programs portably on host systems with different instruction set architectures.
+
+QEMU leverages a Just-In-Time (JIT) compiler called Tiny Code Generator (TCG), which translates guest instructions into host instructions.
+This translation layer is extensible and can be used to add functionality to existing instructions or create new instructions.
+Additionally, the state inside the emulated architecture can be modified to add or remove registers and prepare for functional semantics of potential hardware modifications.
+
+Therefore, the RISC-V implementation in QEMU can be modified to add necessary HFI registers and instructions, allowing for validating HFI's ISA semantics and control-flow traps. 
 Moreover, by implementing in QEMU, RISC-V binaries using the created HFI instructions can be verified for functional correctness.
+This allows the validation of memory safety enforcement, instruction semantics, and sandbox behavior across a variety of test workloads.
 
 % TODO Maybe move below about qemu downsides to future work or limitations section
-Unfortunately, this implementation is not without problems. 
 QEMU lacks support for simulating microarchitectural features such as branch predictors, speculative execution, cache timing, TLB state, and more.
-Therefore, it should be noted that we do implement the serialization on transition instructions to protect against speculative execution outside of the sandbox, an implementation in QEMU will not be able to verify the side-channel resistance.
-Additionally, verfiying performance is problematic as QEMU does not have proper features for verifying cycles.
-
-A more comprehensive implementation for evaluation of security and performance in addition to ISA semantics can be considered in gem5 as done in the original HFI paper \cite{HFI}. 
-However, such an implementation is comparatively much slower to QEMU in modification of the libraries and execution. 
-Therefore, as a feasibility metric, an implementation in QEMU is ideal in our circumstances.
+Moreover, verfiying performance is not possible as QEMU does not have proper features for verifying cycles.
 
@@
 Hence, our work shows the feasibility of HFI in RISC-V and we encourage evaluation similar to that done in the original HFI paper for future work in this area.
+
+\section{Future Work}
+
+\subsection{Cycle Accurate Simulator}
+A more comprehensive implementation for evaluation of security and performance in addition to ISA semantics can be considered in a cycle accurate simulator, such as gem5, as done in the original HFI paper \cite{HFI}.
+In such an implementation, in addition to functional correctness as in QEMU, resistance to spectre attacks can be evaluated. 
+Moreover, with accurate cycle simulator, the performance overhead for HFI implementation can be validated when utilized within a SFI system.
+As an implementation in a cycle accurate simulator is comparatively much slower in modification of the libraries and execution as compared to QEMU, as a feasibility metric for correctness, an implementation in QEMU was ideal with respect to our circumstances.
 

### 2025-04-30 paper.tex
QEMU + related work

@@
 Additionally, the state inside the emulated architecture can be modified to add or remove registers and prepare for functional semantics of potential hardware modifications.
+Furthermore, since these modifications can be applied entirely in software, the need for RTL simulation or FPGA synthesis can be avoided, which would be significantly more complex and time-consuming.
 
-Therefore, the RISC-V implementation in QEMU can be modified to add necessary HFI registers and instructions, allowing for validating HFI's ISA semantics and control-flow traps. 
-Moreover, by implementing in QEMU, RISC-V binaries using the created HFI instructions can be verified for functional correctness.
-This allows the validation of memory safety enforcement, instruction semantics, and sandbox behavior across a variety of test workloads.
-
-% TODO Maybe move below about qemu downsides to future work or limitations section
-QEMU lacks support for simulating microarchitectural features such as branch predictors, speculative execution, cache timing, TLB state, and more.
-Moreover, verfiying performance is not possible as QEMU does not have proper features for verifying cycles.
+While QEMU enables validation of functional correctness, it cannot capture microarchitectural vulnerabilities such as Spectre or Meltdown, which depend on speculative execution, out-of-order pipelines, and cache-based side channels.
+For this reason, QEMU is not suitable for evaluating timing attacks or microarchitectural side effects, but remains a powerful tool for verifying correctness at the architectural level.
+Moreover, verfiying performance is not possible as QEMU lacks features for modeling cycles.
 
@@
 Our approach aims to preserve the security guarantees demanded by modern sandboxing frameworks, while also reducing overheads that historically have hindered fine-grained in-process isolation in the RISC-V ecosystem.
+
+Therefore, the RISC-V implementation in QEMU can be modified to add necessary HFI registers and instructions, allowing for validating HFI's ISA semantics and control-flow traps. 
+Moreover, by implementing in QEMU, RISC-V binaries using the created HFI instructions can be verified for functional correctness.
+This allows the validation of memory safety enforcement, instruction semantics, and sandbox behavior across a variety of test workloads.
 
@@
 Our implementation is able to achieve the sandboxing functionality we wanted and effectively shows that HFI can extend to other architectures beyond x86.
-Hence, our work shows the feasibility of HFI in RISC-V and we encourage evaluation similar to that done in the original HFI paper for future work in this area.
+Hence, our work sufficiently shows the feasibility of HFI in RISC-V.
 
@@
 
+\subsection{Utilize the RISC-V HFI in SFI}
+As previously mentioned, a natural next step is to integrate our RISC-V HFI instructions into an existing SFI compiler to validate the correctness of end-to-end sandboxing behavior. 
+This would involve instrumenting memory accesses and control flow in compiler-generated code and replacing them with HFI-enforced operations. 
+Such an integration would demonstrate the correctness of SFI compilers when relying on hardware enforced isolation mechanisms.
+
+% TODO FIX this maybe if u explain earlier
+Additionally, supporting integration of HFI instructions into existing SFI compilers requires certain specification-focused instructions that we did not implement, such as:
+
+\begin{itemize}
+  \item \texttt{hfi\_get\_version}
+  \item \texttt{hfi\_get\_linear\_code\_range\_count}
+  \item \texttt{hfi\_get\_linear\_data\_range\_count}
+  \item \texttt{hfi\_get\_exit\_reason}
+  \item \texttt{hfi\_get\_exit\_location}
+  \item \textit{and more...}
+\end{itemize}
+
+These additional instructions can be added with minimal modification to the \texttt{funct7} field.
+
 \subsection{Cycle Accurate Simulator}
 A more comprehensive implementation for evaluation of security and performance in addition to ISA semantics can be considered in a cycle accurate simulator, such as gem5, as done in the original HFI paper \cite{HFI}.
-In such an implementation, in addition to functional correctness as in QEMU, resistance to spectre attacks can be evaluated. 
+Since we used QEMU, an environment where the serialization of \textit{hfi\_enter} and \textit{hfi\_exit} would not produce observable effects, we chose not to implement serialization for Spectre mitigation.
+In contrast, a cycle-accurate simulator enables observation of microarchitectural behavior, allowing for both functional correctness validation and evaluation of speculative execution resistance.
 Moreover, with accurate cycle simulator, the performance overhead for HFI implementation can be validated when utilized within a SFI system.
-As an implementation in a cycle accurate simulator is comparatively much slower in modification of the libraries and execution as compared to QEMU, as a feasibility metric for correctness, an implementation in QEMU was ideal with respect to our circumstances.
+As an implementation in a cycle accurate simulator is comparatively much slower in code modification and execution as compared to QEMU, as a feasibility metric for correctness, an implementation in QEMU was ideal with respect to our circumstances.
 

### 2025-05-01 paper.tex
aligned accesses

@@
 
+\usepackage{graphicx}
+
 % *** PDF, URL AND HYPERLINK PACKAGES ***
@@
 One obvious limitation is how we can scale this to work for multiple data regions, which can be a challenge due to the opcode limitation discussed in section 4.
-However, based on the feasibility of our implementation for the one data region and overall structure of the implementation, we believe our proposed ideas from section 4 can allow easy extension to support more explicit data regions.\\
+However, based on the feasibility of our implementation for the one data region and overall structure of the implementation, we believe proposed ideas in section 6 can allow easy extension to support more explicit data regions.\\
 For these new load instructions, we used the free opcode $0001011$.
@@
 \section{Future Work}
+
+\begin{table}[b]
+  \centering
+  \caption{Instruction encoding for \texttt{lw} in RISC-V}
+  {\fontsize{6.5}{7.5}\selectfont
+  \begin{tabular}{|c|c|c|c|c|}
+  \hline
+  \textbf{31--20} & \textbf{19--15} & \textbf{funct3} & \textbf{11--7} & \textbf{OPCODE} \\
+  \hline
+  \texttt{imm[11:0]} & \texttt{rs1} & \texttt{010} & \texttt{rd} & \texttt{0000011} \\
+  \hline
+  \end{tabular}
+  }
+  \label{tab:lw-encoding}
+  \end{table}
+
+  \begin{table}[b]
+  \centering
+  \caption{Instruction encoding for \texttt{sw} in RISC-V}
+  {\fontsize{6.5}{7.5}\selectfont
+  \begin{tabular}{|c|c|c|c|c|c|}
+  \hline
+  \textbf{31--25} & \textbf{24--20} & \textbf{19--15} & \textbf{funct3} & \textbf{11--7} & \textbf{OPCODE} \\
+  \hline
+  \texttt{imm[11:5]} & \texttt{rs2} & \texttt{rs1} & \texttt{010} & \texttt{imm[4:0]} & \texttt{0100011} \\
+  \hline
+  \end{tabular} 
+  }
+  \label{tab:sw-encoding}
+  \end{table}
+
+\subsection{Expand Number of Explicit Regions}
+Previously mentioned, we currently implemented only a singular explict region with regards to the number of instructions that we would introduce.
+In the simplest of implementations, we would just associate a separate instruction for each explicit region, but as we have discussed in section 4, we only have a finite number of opcodes.
+
+Therefore, we need to fit the targeted explicit region number within the load and store instructions we create.
+However, such an observation introduces a larger problem. 
+At closer inspection of the load and store instructions, which are the same as the \texttt{lw} and \texttt{sw} with varying funct3 values, it can be seen that these instructions are extremely tightly packed.
+Thus, any attempt at adding region information will remove information from some other component of the instruction.
+For support of two explicit regions, this will require just 1 bit. Likewise, for supporting four explicit regions as in the original HFI \cite{HFI} paper, we will need 2 bits. 
+As a result, we propose two potential solutions in pursuit of supporting four explicit regions.
+
+Aligning Accesses: 
+In this case, the 2 bits can be sourced from the immediate with the specification that the added immediate will now have to be a multiple of 4. 
+This works without any worries for the loading and storing of a word (4 bytes) and sizes greater than or equal to 4 bytes, but creates issues when considering half words and byte level modifications.
+One possible modification is to also align the halfword accesses to multiples of 2 and have two separate instructions. However, for byte level accesses this is not possible.
+Although this implementation can exclude existing programs that might want to be sandboxed and leverage HFI, in general most code generated from programming languages will be aligned for improved memory access.
+Moreover, any problems that do occur can be solved via compiler modification to modify the unaligned memory accesses to use a combination of byte, halfword, and word versions of the hfi instructions, although this requires substantially more effort.
+
+Reducing Available Registers:
 

### 2025-05-01 paper.tex
Future Work

@@
 
+% TODO: Maybe move
 \begin{table}[b]
@@
 
-Aligning Accesses: 
-In this case, the 2 bits can be sourced from the immediate with the specification that the added immediate will now have to be a multiple of 4. 
-This works without any worries for the loading and storing of a word (4 bytes) and sizes greater than or equal to 4 bytes, but creates issues when considering half words and byte level modifications.
+\textbf{Aligning Accesses}: 
+In this method, the 2 bits can be sourced from the immediate with the specification that the added immediate will now have to be a multiple of 4. 
+This works without any problems for the loading and storing of a word (4 bytes) and sizes greater than or equal to 4 bytes, but creates issues when considering half words and byte level modifications.
 One possible modification is to also align the halfword accesses to multiples of 2 and have two separate instructions. However, for byte level accesses this is not possible.
@@
 
-Reducing Available Registers:
+\textbf{Reducing Available Registers}:
+Here, the 2 bits can be sourced from the indexing of the two registers in the relevant instruction. 
+Registers are input into instructions via a 5 bit value as there are 32 registers in total.
+Therefore, as each of these instructions have two registers in some form (two sources or one source and one destination), 
+we can take a single bit from each of the 5 bits denoting the two registers. 
+This gives us the 2 required bits at the tradeoff of restricting the number of available registers.
+The range of the available registers can additionally be offset or mapped according to a defined set. 
+The use of this method is potentially viable, but more evaluation should be done to verify how problematic register spilling might become.
+
+In reality, a combination of these two methods will probably yield a better result for saving the number of added instructions for new explicit regions.
+Additionally, further evaluation could explore encoding strategies and compiler transformations to better support multiple explicit regions without excessive instruction duplication.
 

### 2025-05-01 paper.tex
h-prefix
Cleanup paragraph formatting

@@
 We began by taking time to understand how QEMU works, the purpose of the various repositories, and gaining a better understanding of RISC-V.
-We worked with the assumption that the system being run on always supports HFI but it is an easy addition to add any necessary hardware checks to validate that assumption.\\
+We worked with the assumption that the system being run on always supports HFI but it is an easy addition to add any necessary hardware checks to validate that assumption.
+
 Our work then began by creating a basic internal register called \textit{hfi\_status}, which serves as a flag bit for verifying whether HFI can be toggled on or off via our newly implemented instructions.
@@
 We are then able to differentiate between the two functions by modifying the funct3 value of the instructions such that \textit{hfi\_enter} has funct3 $= 000$ and \textit{hfi\_enter} has funct3 $= 001$.
-Adding the control and status register (CSR) \textit{hfi\_status} and implementing the \textit{hfi\_enter} and \textit{hfi\_exit} gave us basic functionality to enable and disable HFI.\\
+Adding the control and status register (CSR) \textit{hfi\_status} and implementing the \textit{hfi\_enter} and \textit{hfi\_exit} gave us basic functionality to enable and disable HFI.
 
@@
 We also tested using varying numbers of each region and ensured the HFI logic worked seamlessly with the defined number of regions.
-Implicit regions are defined by a base and a mask.\\
+Implicit regions are defined by a base and a mask.
+
 To store metadata about these regions, we created new structs \textit{HFIImplicitDataRegion} and \textit{HFIImplicitCodeRegion} which we stored in the CPU architecture state in arrays of length based on the number of regions we were supporting.
@@
 Implicit data regions store the enabled bit at bit 7, read bit at bit 6, and write bit at bit 5.
-Implicit code regions store the enabled bit at bit 7 and executable bit at bit 6.\\
+Implicit code regions store the enabled bit at bit 7 and executable bit at bit 6.
+
 Although implicit regions are used for unmodified native binaries, we are still working with the assumption that the software setting up HFI for the binary is a trustable user who will setup the regions properly.
@@
 This was an issue as such a region allowed any code/data addresses to pass through.
-Hence, to prevent this, we set the mask on implicit regions to $0x11111111$ initially which would prevent regions not set up properly to fail HFI checks.\\
+Hence, to prevent this, we set the mask on implicit regions to $0x11111111$ initially which would prevent regions not set up properly to fail HFI checks.
+
 To implement our instruction we have opted to use the free opcode $0001011$.
@@
 This difference is due to the ISA differences between x86 and RISC-V.
-Hence, for our implementation, we focused on providing support for a single region and implemented the following instructions: \textit{hlb0}, \textit{hlh0}, \textit{hlw0}, \textit{hlbu0}, \textit{hlhu0}, \textit{hlwu0}, \textit{hld0}, \textit{hsb0}, \textit{hsh0}, \textit{hsw0}, and \textit{hsd0}.\\
+Hence, for our implementation, we focused on providing support for a single region and implemented the following instructions: \textit{hlb0}, \textit{hlh0}, \textit{hlw0}, \textit{hlbu0}, \textit{hlhu0}, \textit{hlwu0}, \textit{hld0}, \textit{hsb0}, \textit{hsh0}, \textit{hsw0}, and \textit{hsd0}.
+
 For each of the load and store instructions, we would check the whether both the beginning of the access address and end of the address access fall within a given data region (region 0 for the h-prefix and 0-postfix instructions).
 If HFI is enabled and the load or store fall within a properly configured region, then the load or store will be allowed through and proceed to load or store the given address by passing on the access addresses to the corresponding non-HFI load or store instruction.
-If the check fails, HFI would trap and prevent access to the address.\\
+If the check fails, HFI would trap and prevent access to the address.
+
 Explicit data regions are configured using a base and a bound.
@@
 The value passed for a mask in implicit regions was instead used as the bound value when an explicit region was configured.
-For the permissions, explicit data regions store the enabled bit at bit 7, read bit at bit 6, write bit at bit 5, and large region bit at bit 4.\\
+For the permissions, explicit data regions store the enabled bit at bit 7, read bit at bit 6, write bit at bit 5, and large region bit at bit 4.
+
 We tested the implementation using this single explicit data region.
 One obvious limitation is how we can scale this to work for multiple data regions, which can be a challenge due to the opcode limitation discussed in section 4.
-However, based on the feasibility of our implementation for the one data region and overall structure of the implementation, we believe proposed ideas in section 6 can allow easy extension to support more explicit data regions.\\
+However, based on the feasibility of our implementation for the one data region and overall structure of the implementation, we believe proposed ideas in section 6 can allow easy extension to support more explicit data regions.
+
 For these new load instructions, we used the free opcode $0001011$.
@@
 \textit{hfi\_set\_region\_permissions}   &  0001011  &  011     \\ \hline
-\textit{hfi\_hlb0}                       &  0101011  &  000     \\ \hline
-\textit{hfi\_hlh0}                       &  0101011  &  001     \\ \hline
-\textit{hfi\_hlw0}                       &  0101011  &  010     \\ \hline
-\textit{hfi\_hlbu0}                      &  0101011  &  100     \\ \hline
-\textit{hfi\_hlhu0}                      &  0101011  &  101     \\ \hline
-\textit{hfi\_hlwu0}                      &  0101011  &  110     \\ \hline
-\textit{hfi\_hld0}                       &  0101011  &  011     \\ \hline
-\textit{hfi\_hsb0}                       &  1011011  &  010     \\ \hline
-\textit{hfi\_hsh0}                       &  1011011  &  011     \\ \hline
-\textit{hfi\_hsw0}                       &  1011011  &  100     \\ \hline
-\textit{hfi\_hsd0}                       &  1011011  &  000     \\ \hline
+\textit{hlb0}                       &  0101011  &  000     \\ \hline
+\textit{hlh0}                       &  0101011  &  001     \\ \hline
+\textit{hlw0}                       &  0101011  &  010     \\ \hline
+\textit{hlbu0}                      &  0101011  &  100     \\ \hline
+\textit{hlhu0}                      &  0101011  &  101     \\ \hline
+\textit{hlwu0}                      &  0101011  &  110     \\ \hline
+\textit{hld0}                       &  0101011  &  011     \\ \hline
+\textit{hsb0}                       &  1011011  &  010     \\ \hline
+\textit{hsh0}                       &  1011011  &  011     \\ \hline
+\textit{hsw0}                       &  1011011  &  100     \\ \hline
+\textit{hsd0}                       &  1011011  &  000     \\ \hline
 \end{tabular}
@@
 Our testing framework consists of a suite of assembly programs written specifically to test different aspects of HFI functionality on RISC-V.
-These tests are compiled using the RISC-V GCC toolchain and executed on our modified QEMU binary.\\
+These tests are compiled using the RISC-V GCC toolchain and executed on our modified QEMU binary.
+
 We use a modular testing approach where each assembly program focuses on testing a specific component or feature of HFI.
 Our test suite covers several key areas of functionality: basic HFI operations that verify the core instructions function correctly and properly toggle the HFI status; region configuration tests that ensure memory regions can be properly defined with appropriate size and permissions; memory access tests that verify operations within configured regions succeed while operations outside them fail appropriately; and explicit region access tests that exercise the h-prefix load/store instructions to validate boundary checking and permission enforcement.
-We also implemented control tests to establish baseline functionality in non-HFI contexts.\\
+We also implemented control tests to establish baseline functionality in non-HFI contexts.
+
 These test programs are written in RISC-V assembly and leverage a custom set of macros defined in our includes directory that encode our HFI instructions using the appropriate opcodes and operands.
 For example, our HFI enter macro encodes the instruction with opcode 0x0B and function code 0 along with the specified register operands.
-This approach allows us to write readable assembly that directly tests our custom instructions.\\
+This approach allows us to write readable assembly that directly tests our custom instructions.
+
 We developed a Makefile-based build system that compiles these assembly programs using the RISC-V toolchain (riscv64-unknown-elf-as and riscv64-unknown-elf-ld).
-The resulting ELF binaries are then executed on our custom-built QEMU with additional debug flags to capture instruction execution and CPU state information.\\
+The resulting ELF binaries are then executed on our custom-built QEMU with additional debug flags to capture instruction execution and CPU state information.
+
 The test execution outputs are captured in log files, allowing us to analyze instruction execution, register state changes, and any traps or exceptions that occur during testing.
-This approach enables us to verify several critical aspects of the implementation: the HFI instructions execute with the expected opcodes, region configurations are properly stored in CPU state, memory accesses are correctly validated against region bounds, and violations trigger appropriate traps when they should occur.\\
+This approach enables us to verify several critical aspects of the implementation: the HFI instructions execute with the expected opcodes, region configurations are properly stored in CPU state, memory accesses are correctly validated against region bounds, and violations trigger appropriate traps when they should occur.
+
 Through this testing infrastructure, we confirmed the correctness of our implementation across various scenarios including successful sandboxing, proper isolation of memory regions, and correct handling of boundary conditions.
@@
 Our work on HFI for RSIC-V builds on prior work that established HFI for x86-64 architectures \cite{HFI}.
-HFI provides a simple ISA extension to support secure, flexible, and efficient in-process isolation, addressing limitations of existing software-based techniques, including runtime overheads, scalability limits, Spectre vulnerability, and compatibility issues.\\
+HFI provides a simple ISA extension to support secure, flexible, and efficient in-process isolation, addressing limitations of existing software-based techniques, including runtime overheads, scalability limits, Spectre vulnerability, and compatibility issues.
+
 There are also several alternate approaches have been developed over the years for fine-grained isolation involving hardware-assisted techniques.
@@
 Donky facilitates fast domain switches entirely in userspace, with minimal hardware additions on RISC-V.
-Donky has been implemented for both x86 (using MPK emulation) and RISC-V architectures, while IMIX has only been implemented on x86.\\
+Donky has been implemented for both x86 (using MPK emulation) and RISC-V architectures, while IMIX has only been implemented on x86.
+
 IMIX (In-Process Memory Isolation Extension) introduces a minimal x86 ISA extension for isolating sensitive data within a process \cite{IMIX}.
@@
 IMIX was prototyped using Intel’s simulation infrastructure, with extensions to LLVM and Linux to support its new instruction semantics.
-The authors demonstrated that IMIX significantly reduces the attack surface for in-process defenses with low overhead and better flexibility than existing solutions like MPK, which are limited by unprivileged key registers and require complex mitigation workarounds to be secure.\\
+The authors demonstrated that IMIX significantly reduces the attack surface for in-process defenses with low overhead and better flexibility than existing solutions like MPK, which are limited by unprivileged key registers and require complex mitigation workarounds to be secure.
+
 In addition to this, there are also methods that make use of capability-based addressing, a scheme for controlling access to memory, which is seen most prominently in CHERI \cite{CHERI}.
@@
 However, adapting and recompiling application code is generally required to utilize its full protection features, and achieving comprehensive object-level memory safety involves extensive modifications throughout the software stack (OS, ABI, compiler).
-CHERI offers strong fine-grained protection within an address space and supports scalable userspace protection domains.\\
+CHERI offers strong fine-grained protection within an address space and supports scalable userspace protection domains.
+
 In summary, while Donky leverages protection keys and a userspace monitor with hardware protection for the policy register, and CHERI introduces capabilities for fine-grained, unforgeable memory references requiring significant software stack changes, IMIX proposes a lightweight ISA extension with minimal hardware cost to securely restrict access to sensitive memory used by diverse in-process defenses.

### 2025-05-01 paper.tex
Encoding Instructions

@@
 
+\usepackage{float}
+
 % *** Do not adjust lengths that control margins, column widths, etc. ***
@@
 \section{Approach}
+\label{sec:approach}
+
 Our work brings HFI concepts to the RISC-V ecosystem by adapting the existing HFI design while fitting into RISC-V's existing privileged architecture. 
@@
 
+\subsection{Custom Instruction Encoding in RISC-V}
+Before we consider custom instructions, we will first consider the key details of instruction encoding in RISC-V.
+For starters, RISC-V is a fixed instruction length architecture with two variant sizes 16-bit (compressed) and 32-bit.
+For our purposes, we will only need to consider the 32-bit instructions.
+
+% TODO put in better spot
+\begin{table}[H]
+  \centering
+  \caption{RISC-V Instruction Format Encodings}
+  {\fontsize{6.5}{7.5}\selectfont
+  \begin{tabular}{|c|c|c|c|c|c|c|}
+  \hline
+  \textbf{31--25} & \textbf{24--20} & \textbf{19--15} & \textbf{14--12} & \textbf{11--7} & \textbf{6--0} & \textbf{Type} \\
+  \hline
+  \texttt{funct7} & \texttt{rs2} & \texttt{rs1} & \texttt{funct3} & \texttt{rd} & \texttt{opcode} & R-type \\
+  \hline
+  \multicolumn{2}{|c|}{\texttt{imm[11:0]}} & \texttt{rs1} & \texttt{funct3} & \texttt{rd} & \texttt{opcode} & I-type \\
+  \hline
+  \texttt{imm[11:5]} & \texttt{rs2} & \texttt{rs1} & \texttt{funct3} & \texttt{imm[4:0]} & \texttt{opcode} & S-type \\
+  \hline
+  \end{tabular} 
+  }
+  \label{tab:encoding-summary}
+\end{table}
+
+Although Table~\ref{tab:encoding-summary} is not a complete representation of all the types of instructions in RISC-V, it will suffice for our implementation.
+Register-type (R-type) denotes instructions that function entirely on registers such as \textit{add} and \textit{or}.
+Immediate-type (I-type) are instructions that will require an immediate such as \textit{lw} (load word) and \textit{addi}.
+Store-type (S-type) are specified for specifically stores to memory where the immediate is split into two parts; an instruction of this variant is \textit{sw} (store word).
+For our purposes R-type will be utilized for updating and retrieving state of the sandbox such as the size of the implicit regions, base and bound of explicit regions, and their associated permissions.
+I-type and S-type will be in regards to the loads and stores for the explicit regions with the h-prefixed instructions mentioned in Section~\ref{sec:impl-eval}.
+
+In each of these instruction types, certain fields remain consistent, these are the first source register (rs1), the funct3 field, and the opcode. 
+The funct3 and opcode fields are used together to distinguish between specific instructions. 
+The opcode generally defines the broad category of the instruction, while funct3 provides finer-grained differentiation within that category.
+
+\begin{table}[H]
+  \centering
+  \caption{Encoding of RISC-V Load Instructions}
+  \small
+  {
+  \begin{tabular}{|c|c|c|c|c|c|}
+    \hline
+    \textbf{imm} & \textbf{rs1} & \textbf{funct3} & \textbf{rd} & \textbf{opcode} & \textbf{Instruction} \\
+    \hline
+    \texttt{imm[11:0]} & \texttt{rs1} & \texttt{000} & \texttt{rd} & \texttt{0000011} & \texttt{lb} \\
+    \texttt{imm[11:0]} & \texttt{rs1} & \texttt{001} & \texttt{rd} & \texttt{0000011} & \texttt{lh} \\
+    \texttt{imm[11:0]} & \texttt{rs1} & \texttt{010} & \texttt{rd} & \texttt{0000011} & \texttt{lw} \\
+    \texttt{imm[11:0]} & \texttt{rs1} & \texttt{100} & \texttt{rd} & \texttt{0000011} & \texttt{lbu} \\
+    \texttt{imm[11:0]} & \texttt{rs1} & \texttt{101} & \texttt{rd} & \texttt{0000011} & \texttt{lhu} \\
+    \texttt{imm[11:0]} & \texttt{rs1} & \texttt{110} & \texttt{rd} & \texttt{0000011} & \texttt{lwu} \\
+    \texttt{imm[11:0]} & \texttt{rs1} & \texttt{011} & \texttt{rd} & \texttt{0000011} & \texttt{ld} \\
+    \hline
+  \end{tabular}
+  }
+  \label{tab:load-instructions}
+\end{table}
+
+For example, in Table~\ref{tab:load-instructions}, see that their opcode is the same indicating their broad category of load and the funct3 value is varied for the specialized instruction.
+
+Currently, there are 4 opcodes that are available for extending the instruction set; these are custom-0 (0001011), custom-1 (0101011), custom-2 (1011011), and custom-3 (1111011).
+It is important to note that the opcodes custom-2 and custom-3 have been reserved for the open proposal for RV128, which will specify the RISC-V ISA for 128-bit machines. 
+However, these remain available in RV32 and RV64. 
+For simplicity, we will focus our efforts on extending RV64 to ensure no potential conflicts, but modifications should be considered to reduce the number of used opcodes and instructions to maximize compatibility.
+
+In the composition of new instructions that interact with the underlying HFI state using the R-type instructions, we will use the custom-0 instruction.
+Additionally, if we return to Table~\ref{tab:encoding-summary}, we can see that these R-type instructions also have a funct7 value which can be used to further increase the number of instructions with a fixed value for funct3 and opcode.
+Therefore, further consolidation can be performed for many of the R-type instructions. 
+However, it should be mentioned that in creating instructions to expand functionality such as in Section~\ref{sec:impl-eval}, we did not use this as a focus, and determined this could be determined later.
+As a result, our chosen funct3 and funct7 values can be consolidated within many of our implemented instructions.
+
+Furthermore, it can be seen that loads and stores have little room for expansion with the absence of a funct7 value, and for the explicit region problems arise, which are discussed in Section~\ref{sec:future-work}.
+
+\subsection{Extending QEMU for HFI}
+
+\subsection{Region Enforcement Semantics}
+
 % TODO add the opcode limitation for explicit regions and possible solutions
@@
 \section{Implementation \& Evaluation}
+\label{sec:impl-eval}
+
 \subsection{Implementation - Intial Setup}
@@
 We tested the implementation using this single explicit data region.
-One obvious limitation is how we can scale this to work for multiple data regions, which can be a challenge due to the opcode limitation discussed in section 4.
-However, based on the feasibility of our implementation for the one data region and overall structure of the implementation, we believe proposed ideas in section 6 can allow easy extension to support more explicit data regions.
+One obvious limitation is how we can scale this to work for multiple data regions, which can be a challenge due to the opcode limitation discussed in Section~\ref{sec:approach}.
+However, based on the feasibility of our implementation for the one data region and overall structure of the implementation, we believe proposed ideas in Section~\ref{sec:future-work} can allow easy extension to support more explicit data regions.
 
@@
 \section{Future Work}
+\label{sec:future-work}
+
+\subsection{Expand Number of Explicit Regions}
+Previously mentioned, we currently implemented only a singular explict region with regards to the number of instructions that we would introduce.
+In the simplest of implementations, we would just associate a separate instruction for each explicit region, but as we have discussed in Section~\ref{sec:approach}, we only have a finite number of available opcodes.
 
 % TODO: Maybe move
-\begin{table}[b]
+\begin{table}[H]
   \centering
@@
   \hline
-  \textbf{31--20} & \textbf{19--15} & \textbf{funct3} & \textbf{11--7} & \textbf{OPCODE} \\
+  \textbf{31--20} & \textbf{19--15} & \textbf{funct3} & \textbf{11--7} & \textbf{opcode} \\
   \hline
@@
 
-  \begin{table}[b]
+  \begin{table}[H]
   \centering
@@
   \hline
-  \textbf{31--25} & \textbf{24--20} & \textbf{19--15} & \textbf{funct3} & \textbf{11--7} & \textbf{OPCODE} \\
+  \textbf{31--25} & \textbf{24--20} & \textbf{19--15} & \textbf{funct3} & \textbf{11--7} & \textbf{opcode} \\
   \hline
@@
   \end{table}
-
-\subsection{Expand Number of Explicit Regions}
-Previously mentioned, we currently implemented only a singular explict region with regards to the number of instructions that we would introduce.
-In the simplest of implementations, we would just associate a separate instruction for each explicit region, but as we have discussed in section 4, we only have a finite number of opcodes.
 
@@
 
+\bibitem{RISCVSpec}
+“V Ratified Specifications.” RISC, riscv.org/specifications/ratified/. Accessed 30 Apr. 2025.
+
 \end{thebibliography}

### 2025-05-01 paper.tex
Approach

@@
 
-\subsection{Extending QEMU for HFI}
+\subsection{Extending QEMU for HFI: Instructions}
+Upon determining the instruction encoding, decoding the instruction requires modifiying the translation pipeline to recognize the instruction and translate it into the appropriate sequence of TCG operations for execution.
+This process is structured into three main components:
+
+First, a decode rule must be added in the appropriate instruction format file (e.g., \texttt{insn32.decode}) to match the opcode and instruction fields.
+This rule identifies the custom instruction and extracts its parameters (such as \texttt{rd}, \texttt{rs1}, \texttt{imm}, etc.).
+
+Second, a corresponding translation function must be defined in the instruction translation file (e.g., \texttt{trans\_hfi.c.inc}).
+This function is written using QEMU's Tiny Code Generator (TCG) and is responsible for generating the intermediate representation that will be executed by the emulator.
+
+Finally, a helper function may be defined (typically in C) to carry out the instruction's semantic behavior on the CPU.
+The translation function emits a call to this helper using TCG, and passes in the appropriate virtual CPU state and arguments.
+This separation allows the architectural logic to be written in C while still integrated into JIT-compiled TCG execution.
+
+This process forms the standard method for implementing new instructions in QEMU, including those used in supporting architectural extensions like HFI.
+
+\subsection{Extending QEMU for HFI: State}
+Modifying the state of QEMU to support the state required for HFI involves extending the architectural state of the RISC-V CPU model to track region metadata and HFI status.
+
+All state must be added to QEMU's \texttt{CPURISCVState} structure, which holds per-core architectural state.
+This enables extensions such as HFI to be modeled as a first-class architectural feature that can interact with instruction translation and memory access logic.
+
+The added new state is zeroed out on initialization, but modifications are preserved across QEMU execution so that memory checks and instruction semantics can depend on it.
 
 \subsection{Region Enforcement Semantics}
+Region enforcement in QEMU for HFI can be achieved by emitting TCG-level checks that guard memory accesses when HFI is active.
+These checks validate that each memory access falls within the appropriate region bounds, either for implicit data/code regions or explicit ones depending on the instruction type.
+
+For standard load and store instructions that operate under implicit region constraints, the necessary bounds checks must be inserted during translation.
+Specifically, TCG code should be emitted to verify that the computed memory address lies within a configured implicit region before allowing the memory operation to proceed.
+These checks must always be emitted unconditionally, because QEMU caches translated blocks.
+If the TCG check were omitted based on the runtime value of the HFI status, the translation block could be cached with a stale value, leading to incorrect execution that bypasses necessary HFI enforcement.
+
+In contrast, HFI-prefixed load and store instructions perform region validation directly in their corresponding C helper functions, which are invoked from TCG.
+These helpers are written to emit the necessary checks using standard C logic, and since the code is incorporated into TCG, the effect is functionally equivalent to always emitting the inline TCG code as with implcit regions.
+
+Together, these approaches ensure that region enforcement is always active when required and that both implicit and explicit region checks are performed correctly and reliably across QEMU's translated execution model.
 

### 2025-05-01 paper.tex
add to project goals

@@
 Hardware-assisted fault isolation (HFI) on RISC-V aims to address two critical needs in modern computing systems: preventing security bugs by isolating untrusted or risky code, and better understanding (and improving) performance overheads that arise from current software-based or OS-based sandboxing mechanisms.
+Fine-grained sandboxing is essential here as a single defective or malicious component can be contained before it compromises the rest of the application or operating system.
 By offering a minimal set of architectural extensions for in-process isolation, HFI empowers developers to sandbox components (e.g., libraries, plugins, user-defined functions) without resorting to heavyweight processes or virtual machines.
@@
 Several recent high-profile vulnerabilities—such as Heartbleed (CVE-2014-0160) and Spectre (CVE-2017-5753, CVE-2017-5715)—underscore the need for stronger isolation between mutually distrusting code modules.
+In each of these incidents, the lack of strong intra-process boundaries allows an attacker to read or corrupt sensitive data; a robust sandbox would have limited the impact of the attack to the faulty module itself and preserved overall system integrity.
+Compared with classic software-based fault isolation (SFI), a hardware sandbox not only reduces runtime overhead but also eliminates whole classes of security issues that have historically plagued SFI implementations such as side-channel attacks and out-of-bounds errors.
 System architects and software developers alike have turned to hardware features for efficient compartmentalization.

### 2025-05-01 paper.tex
added CVE info to security section

@@
 
+Equally important, HFI tackles concrete flaws that consistently slip past software guards.
+Software-only SFI cannot stop transient-execution attacks such as Spectre (CVE-2017-5753, CVE-2017-5715): an attacker can speculatively jump over the compiler-inserted bounds checks and leak data.
+Because HFI performs the range comparison inside the load/store pipeline and stalls the access until the check resolves, the classic Spectre gadgets are rendered unexploitable.
+The same in-pipeline enforcement also neutralises newer variants (e.g., Spectre v4 and Load Value Injection) without utilizing costly lfences throughout the code base.
+
+At the same time, HFI preserves the ability to confine more “traditional” memory-safety bugs, such as out-of-bounds errors, that SFI already blocks—while doing so with better performance.
+Isolating OpenSSL in an HFI sandbox still prevents the Heartbleed over-read (CVE-2014-0160) from escaping into adjacent heaps, but avoids the 25–35\% overhead that guard-page faults imposed in process-based SFI.
+Likewise, keeping vulnerable image decoders (e.g., the 2023 libwebp heap overflow) in an HFI region lets a browser switch into and out of the decoder in ~80 ns—about the cost of a function call—rather than launching a heavyweight renderer process.
+In short, HFI simultaneously closes the speculative side-channels that SFI misses and slashes the overhead of the mitigations SFI already provides.
+
 \section{Approach}

### 2025-05-01 paper.tex
Updated Paper

@@
 \subsection{Implementation - Intial Setup}
-Since we were implenting our RISC-V HFI extension in QEMU, we cloned the current open-source QEMU repository and began work there.
+Since we were implementing our RISC-V HFI extension in QEMU, we cloned the current open-source QEMU repository and began work there.
 We began by taking time to understand how QEMU works, the purpose of the various repositories, and gaining a better understanding of RISC-V.

### 2025-05-01 paper.tex
Updated Paper

@@
 Traditionally, general isolation was considered at the granularity of processes, containers, and virtual machines, which do achieve the functionality of isolation of libraries through hardware-level protection via the MMU and user-kernel privileges.
-Although such these isolation methods are functional, they have significant performance overhead from the context switching and memory management. 
+Although these isolation methods are functional, they have significant performance overhead from the context switching and memory management. 
 Due to the fact that the executing libraries are quite small, it motivates isolation that is fine-grained and lightweight.
@@
 In order to implement our instruction we have opted to use the free opcode $0001011$.
-We are then able to differentiate between the two functions by modifying the funct3 value of the instructions such that \textit{hfi\_enter} has funct3 $= 000$ and \textit{hfi\_enter} has funct3 $= 001$.
+We are then able to differentiate between the two functions by modifying the funct3 value of the instructions such that \textit{hfi\_enter} has funct3 $= 000$ and \textit{hfi\_exit} has funct3 $= 001$.
 Adding the control and status register (CSR) \textit{hfi\_status} and implementing the \textit{hfi\_enter} and \textit{hfi\_exit} gave us basic functionality to enable and disable HFI.
@@
 \subsection{Implementation - Explicit Regions}
-Next, we moved on to implement explicit regions by defining a new struct \textit{HFIExpllicitDataRegion} to store metadata about explicit regions.
+Next, we moved on to implement explicit regions by defining a new struct \textit{HFIExplicitDataRegion} to store metadata about explicit regions.
 These structs were also stored in a new array in the CPU Architecture State of length based on the global variable indicating the number of regions we wanted to support.
@@
 
-For each of the load and store instructions, we would check the whether both the beginning of the access address and end of the address access fall within a given data region (region 0 for the h-prefix and 0-postfix instructions).
+For each of the load and store instructions, we would check whether both the beginning of the access address and end of the address access fall within a given data region (region 0 for the h-prefix and 0-postfix instructions).
 If HFI is enabled and the load or store fall within a properly configured region, then the load or store will be allowed through and proceed to load or store the given address by passing on the access addresses to the corresponding non-HFI load or store instruction.
@@
 The wasm2c code converts wasm files to C source and header and we wanted to modify it to use our modified h-prefix load and store instructions.
-We were unable to complete this step unfortunately due to some issues with building and running the module.
+Unfortunately, we were unable to complete this due to some issues with building and running the module.
 However, this doesn't detract from the implementation of HFI itself.
@@
 \subsection{Expand Number of Explicit Regions}
-Previously mentioned, we currently implemented only a singular explict region with regards to the number of instructions that we would introduce.
+Previously mentioned, we currently implemented only a singular explicit region with regards to the number of instructions that we would introduce.
 In the simplest of implementations, we would just associate a separate instruction for each explicit region, but as we have discussed in Section~\ref{sec:approach}, we only have a finite number of available opcodes.
@@
 In contrast, a cycle-accurate simulator enables observation of microarchitectural behavior, allowing for both functional correctness validation and evaluation of speculative execution resistance.
-Moreover, with accurate cycle simulator, the performance overhead for HFI implementation can be validated when utilized within a SFI system.
+Moreover, with an accurate cycle simulator, the performance overhead for HFI implementation can be validated when utilized within a SFI system.
 As an implementation in a cycle accurate simulator is comparatively much slower in code modification and execution as compared to QEMU, as a feasibility metric for correctness, an implementation in QEMU was ideal with respect to our circumstances.

### 2025-05-01 paper.tex
Small grammar and spelling fixes

@@
 For this reason, QEMU is not suitable for evaluating timing attacks or microarchitectural side effects, but remains a powerful tool for verifying correctness at the architectural level.
-Moreover, verfiying performance is not possible as QEMU lacks features for modeling cycles.
+Moreover, verifying performance is not possible as QEMU lacks features for modeling cycles.
 
@@
 Additionally, we have \textit{hfi\_exit}, which sets \textit{hfi\_status} $=0$.
-RISC-V contains unmapped opcodes that can be used for the additional of new instructions.
+RISC-V contains unmapped opcodes that can be used for the addition of new instructions.
 In order to implement our instruction we have opted to use the free opcode $0001011$.
@@
 These structs were also stored in a new array in the CPU Architecture State of length based on the global variable indicating the number of regions we wanted to support.
-While the original HFI paper had the h-prefix \textit{mov} instruct (\textit{hmov}), we had h-prefix load and store instruction.
+While the original HFI paper had the h-prefix \textit{mov} instruction (\textit{hmov}), we had h-prefix load and store instruction.
 This difference is due to the ISA differences between x86 and RISC-V.
@@
 \subsection{Expand Number of Explicit Regions}
-Previously mentioned, we currently implemented only a singular explicit region with regards to the number of instructions that we would introduce.
+Previously mentioned, we have currently implemented only a singular explicit region with regards to the number of instructions that we would introduce.
 In the simplest of implementations, we would just associate a separate instruction for each explicit region, but as we have discussed in Section~\ref{sec:approach}, we only have a finite number of available opcodes.
@@
 \section{Related Work}
-Our work on HFI for RSIC-V builds on prior work that established HFI for x86-64 architectures \cite{HFI}.
+Our work on HFI for RISC-V builds on prior work that established HFI for x86-64 architectures \cite{HFI}.
 HFI provides a simple ISA extension to support secure, flexible, and efficient in-process isolation, addressing limitations of existing software-based techniques, including runtime overheads, scalability limits, Spectre vulnerability, and compatibility issues.
